---
title: "Data Processing"
author: "YoungZhang"
date: "2025-04-01"
output: pdf_document
---

**Logs:**

2025-04-01: I convert the load, humidity, and temperature data into daily and hourly data, then save the csv files to the "processed" file. Note: I noticed that there are 6 missing variables for all the 3 series, I filled the missings using interpolation method. But I didn't do anything for the daily data.

---

```{r}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=60), tidy=TRUE) 
```

```{r package, message=FALSE}
library(readxl)
library(openxlsx)
library(dplyr)
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(smooth)
library(zoo)
library(kableExtra)
```

Importing Data
```{r}
#importing data
load <- read_excel("../data/raw/load.xlsx")   # I changed the to the relative address 

#Creating a data frame with daily observations
load <- load %>% 
  rowwise() %>%
  mutate(daily_avg = mean(c_across(h1:h24), na.rm = TRUE)) %>%
  ungroup()

ggplot(load, aes(x=date,y=daily_avg)) +
  geom_line() +
  ylab("Average Daily Load")

summary(load$daily_avg)

```

Converting into time series
```{r}
load_daily <- msts(load$daily_avg, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,01,01))
```


Load humidity and temperature data
```{r}

## Transform hourly data of humidity and temperature
# Here, I first average the humidity/temperature of 28 stations within one hour. Then I summarize to get daily values


humidity <- read_excel("../data/raw/relative_humidity.xlsx")


humidity_daily <- humidity %>% 
  rowwise() %>%
  mutate(humidity_avg = mean(c_across(rh_ws1:rh_ws28), na.rm = TRUE)) %>%  
  ungroup() %>%
  group_by(date) %>%
  summarize(daily_avg_humidity = mean(humidity_avg, na.rm = TRUE))   


humidity_daily_ts <- msts(humidity_daily$daily_avg_humidity, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,01,01))


## Transform hourly data of temperature

temperature <- read_excel("../data/raw/temperature.xlsx")

temperature_daily <- temperature %>% 
  rowwise() %>%
  mutate(temperature_avg = mean(c_across(t_ws1:t_ws28), na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(date) %>%
  summarize(daily_avg_temperature = mean(temperature_avg, na.rm = TRUE))


temperature_daily_ts <- msts(temperature_daily$daily_avg_temperature, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,01,01))

autoplot(humidity_daily_ts, series = "humidity")+
  autolayer(temperature_daily_ts, series = "temperature")


## Combine the three series together

combined_daily_series <- cbind(load_daily,humidity_daily_ts,temperature_daily_ts)


```


Output processed data
```{r}
## Output the daily data

combined_ts_df <- as.data.frame(combined_daily_series)

write.csv(combined_ts_df, "../data/processed/processed_daily_data.csv", row.names = FALSE)

```

```{r}
## Output hourly data
load <- read_excel("../data/raw/load.xlsx")
load_long <- load %>%
  pivot_longer(cols = starts_with("h"),  
               names_to = "hr",          
               names_prefix = "h",      
               values_to = "load_value") 

load_long$hr <- as.numeric(load_long$hr)


humidity <- read_excel("../data/raw/relative_humidity.xlsx")
humidity <- humidity %>% 
  rowwise() %>%
  mutate(humidity_avg = mean(c_across(rh_ws1:rh_ws28), na.rm = TRUE)) %>%  
  ungroup()



temperature <- read_excel("../data/raw/temperature.xlsx")
temperature <- temperature %>% 
  rowwise() %>%
  mutate(temperature_avg = mean(c_across(t_ws1:t_ws28), na.rm = TRUE)) %>%
  ungroup() 



combined_hourly  <- load_long %>%
  left_join(humidity %>% select(date, hr, humidity_avg), by = c("date", "hr"))%>%
  left_join(temperature %>% select(date, hr, temperature_avg), by = c("date", "hr"))


# deal with missing values of hourly data
combined_hourly$humidity_avg <- na.approx(combined_hourly$humidity_avg)
combined_hourly$temperature_avg <- na.approx(combined_hourly$temperature_avg)
combined_hourly$load_value <- na.approx(combined_hourly$load_value)


write.csv(combined_hourly, "../data/processed/processed_hourly_data.csv", row.names = FALSE)

```



