---
title: "Models"
author: "YoungZhang"
output: html_document
date: "2025-04-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=60), tidy=TRUE) 
```

```{r}
library(readxl)
library(openxlsx)
library(dplyr)
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(smooth)
library(zoo)
library(kableExtra)
```

## Importing Data

```{r pressure, echo=FALSE}

#Importing daily data
daily_data <- read.csv(file="/home/guest/ForecastingCompetition/data/processed/processed_daily_data.csv", header=TRUE)

#Importing hourly data
hourly_data <- read.csv(file="/home/guest/ForecastingCompetition/data/processed/processed_hourly_data.csv", header=TRUE)

#Inspect daily data
head(daily_data)
nvar <- ncol(daily_data) - 1
nobs <- nrow(daily_data)

#Inspect hourly data
head(hourly_data)
nvar <- ncol(hourly_data) - 1
nobs <- nrow(hourly_data)

```
## Testing and Training Datasets - Daily
```{r}

train_daily <- daily_data %>%
  filter(date < as.Date("2010-01-01"))

test_daily <- daily_data %>%
  filter(date >= as.Date("2010-01-01"))

ts_load_daily_u <- msts(daily_data$load_daily, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,01,01))

ts_temp_daily_u <- msts(daily_data$temperature_daily_ts, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,01,01))

ts_humid_daily_u <- msts(daily_data$humidity_daily_ts, 
                           seasonal.periods =c(7,365.25),
                           start=c(2005,01,01))


n <- length(ts_load_daily_u)
h <- 365.25

ts_daily_train_u <- window(ts_load_daily_u, end=c(2009, 365))

ts_daily_test_u <- window(ts_load_daily_u, start=c(2010,1))

```

## Testing and Training Datasets - Hourly
```{r}

ts_load_hourly_u <- msts(hourly_data$load_value, 
                           seasonal.periods =c(24,168,8766),
                           start=c(2005,01,01))

ts_hourly_train_u <- subset(ts_load_hourly_u,
                                   end = length(ts_load_hourly_u)-365)


ts_hourly_test_u <- subset(ts_load_hourly_u,
                                   start = length(ts_load_hourly_u)-365)

```


## Initial Plots & Time Series
```{r}

TSplot_u <- autoplot(ts_load_daily_u) +
    ggtitle("Daily Load Time Series (2005-2010)") +
    xlab("Year") + 
    ylab("Load") +
    theme_minimal()
plot(TSplot_u)

#ACF and PACF plots
par(mfrow=c(1,2))
ACF_Plot_u <- Acf(daily_data$load_daily, lag = 40, plot = TRUE)
PACF_Plot_u <- Pacf(daily_data$load_daily, lag = 40)
par(mfrow=c(1,1))

```

## Forecast using Seasonal Naive Method - Daily
```{r}

# Seasonal naive on original data
SNAIVE_seas <- snaive(ts_load_daily_u, h=60)
plot(SNAIVE_seas)

checkresiduals(SNAIVE_seas)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Write into Excel file
```{r}

write.csv(SNAIVE_seas, file = "seasonal_naive_forecast.csv")

```


## TBATS - Daily
```{r}

tbats_fit <- tbats(ts_load_daily_u)
tbats_forecast <- forecast(tbats_fit, h = 60)
autoplot(tbats_forecast)

```

## Write CSV for tbats
```{r}

write.csv(tbats_forecast, file = "tbats_forecast.csv")

```

## STL + ETS - Daily
```{r}

load_stl_ets <- stlf(ts_load_daily_u, h = 60, method = "ets")
plot(load_stl_ets)

```

## Write CSV for STL + ETS
```{r}

write.csv(load_stl_ets, file = "stlets_forecast.csv")

```

## Arithmetic mean
```{r}

MEAN_seas <- meanf(y = ts_load_daily_u, h = 60)  
checkresiduals(MEAN_seas)
plot(MEAN_seas)

write.csv(MEAN_seas, file = "MEAN_seas_forecast.csv")

```


## Forecast using Seasonal Naive Method - hourly
```{r}

# Seasonal naive on original data
SNAIVE_seash <- snaive(ts_load_hourly_u, h=1440)
plot(SNAIVE_seash)

checkresiduals(SNAIVE_seash)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Write into CSV
```{r}

write.csv(SNAIVE_seash, file = "seasonal_naive_hourly_forecast.csv")

```

## ARIMA + Fourier
```{r}

ts_load_daily_uf <- ts(ts_load_daily_uf, frequency = 365)

K <- 6
fourier_train <- fourier(ts_load_daily_uf, K = K)
fourier_future <- fourier(ts_load_daily_uf, K = K, h = 60)
fit_fourier <- auto.arima(ts_load_daily_uf, xreg = fourier_train, seasonal = FALSE)
fourier <- forecast(fit_fourier, xreg = fourier_future, h = 60)

plot(fourier)

write.csv(fourier, file = "fourier_forecast.csv")


```







